---
title: ""
author:
  - "Aryaman Dhindsa"
  - "Joshua Bannon"
  - "Smit Tambat"
  - "Chunjian Zhang"
course: "ETC5513"
unit: "Collaborative and Reproducible Practices"
semester: "Semester 1, 2026"
date: "27 May 2026"
format: html
self-contained: TRUE
bibliography: reference.bib
---

# Executive Summary

# Introduction

This report examines how the production budgets and box-office revenues of Netflix films changed before, during, and after COVID-19 period. We define three periods: pre-COVID (2016-2019), COVID (2020-2021), and post-COVID (2022-2023). Using a public dataset of Netflix films titles, we compare yearly means of budget and revenue across these periods to identify any pandemic related shifts. The analysis focuses on financial measures rather than release volume, as the dataset is balanced by release year and therefore unsuitable for measuring output. Understanding these patterns offers insight into how the pandemic affected the economics of Netflix film production. Specifically, we ask: **How did the Budgets and Revenues of Netflix films change before, during, and after COVID-19?** 

# Methodology
The analysis draws on a public dataset of 16,000 Netflix film titles, each recorded with its release year and an average audience vote score. We dropped the 2025 titles before doing anything else. Those films had only just been released, most had barely been rated, and leaving them in pulled the averages down to numbers that clearly weren't real. That left 15,000 films spanning 2010 to 2024 to work with.

To get at the pandemic question, we sorted the films into three periods:
pre-COVID (2010–2019), COVID (2019–2021), and post-COVID (2022–2023). For each period we worked out the mean budget, which gives us a simple way to compare how much the revenue changed compared to the budget across the three windows. We also calculated the average rating year by year, so we could see the smaller movements that a three-way split tends to hide. Everything was done in R using the `tidyverse` packages, and the code sits in the same Quarto document as this report so the whole thing can be reproduced from the raw data.

@fig-yearly plots the yearly average rating from 2010 to 2024, with the two pandemic years marked so any shift around them is easy to spot. @tbl-periods then pulls this together, showing the mean budget and mean revenue in each period side by side. Between them, the figure gives the year-to-year detail and the table gives the bigger comparison.

```{r}
#| label: fig-yearly
#| fig-cap: "Yearly mean budget and revenue for Netflix films, 2016–2023 (USD)."
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

netflix <- read_csv("data/netflix_data.csv") |>
  filter(release_year >= 2016, release_year <= 2024,
         budget > 0, revenue > 0)

yearly <- netflix |>
  group_by(release_year) |>
  summarise(
    `Mean budget` = mean(budget, na.rm = TRUE),
    `Mean revenue` = mean(revenue, na.rm = TRUE)
  ) |>
  pivot_longer(cols = c(`Mean budget`, `Mean revenue`),
               names_to = "Measure", values_to = "USD")

ggplot(yearly, aes(x = release_year, y = USD, colour = Measure)) +
  geom_line(linewidth = 1) +
  geom_point() +
  geom_vline(xintercept = c(2019, 2021), linetype = "dashed",
             colour = "grey50") +
  scale_y_continuous(labels = scales::label_dollar(scale = 1e-6,
                                                    suffix = "M")) +
  scale_colour_manual(values = c("Mean budget" = "#221F1F",
                                  "Mean revenue" = "#E50914")) +
  labs(x = "Release year", y = "Mean (USD)", colour = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: tbl-periods
#| tbl-cap: "Mean budget and revenue by COVID-19 period (USD millions)."
#| tbl-cap-location: bottom
#| echo: false
#| message: false
#| warning: false

library(knitr)

netflix |>
  mutate(period = case_when(
    release_year <= 2018 ~ "Pre-COVID (2016–2018)",
    release_year <= 2021 ~ "COVID (2019–2021)",
    TRUE                 ~ "Post-COVID (2022–2024)"
  )) |>
  group_by(period) |>
  summarise(
    `Mean budget (USD M)` = round(mean(budget, na.rm = TRUE) / 1e6, 1),
    `Mean revenue (USD M)` = round(mean(revenue, na.rm = TRUE) / 1e6, 1),
    `Number of films` = n()
  ) |>
  arrange(factor(period, levels = c(
    "Pre-COVID (2016–2018)", "COVID (2019–2021)", "Post-COVID (2022–2024)"
  ))) |>
  rename(Period = period) |>
  kable()
```

# Result
As shown in @fig-yearly, revenue experienced a sharp decline at the beginning of the COVID-19 period, despite only a slight reduction in budget allocation. However, during the peak of the pandemic, revenue increased significantly, which was accompanied by a corresponding rise in Netflix's budget during the same period.

The budget reached approximately $45 million in 2021 and subsequently fluctuated between $25 million and $50 million through 2023. Revenue, on the other hand, increased substantially after 2020 and continued to show steady growth from 2021 onward.

A summarized comparison of the pre-COVID, during-COVID, and post-COVID periods is also presented in @tbl-periods.

```{r}
#| label: tbl-movie-release
#| tbl-cap: "Average no. of Movies released during the three periods"
#| tbl-cap-location: bottom
#| echo: false
#| message: false
#| warning: false

period_summary <- netflix |>
  mutate(
    period = case_when(
      release_year < 2019 ~ "Pre-COVID (2016–2018)",
      release_year %in% 2019:2021 ~ "COVID (2019–2021)",
      release_year > 2021 ~ "Post-COVID (2022–2023)"
    )
  ) |>
  group_by(period, release_year) |>
  summarise(
    movies_released = n(),
    .groups = "drop"
  ) |>
  group_by(period) |>
  summarise(
    `Average movies released per year` = mean(movies_released)
  )

period_summary |>
  arrange(factor(period, levels = c(
    "Pre-COVID (2016–2018)", "COVID (2019–2021)", "Post-COVID (2022–2023)"
  ))) |>
  rename(Period = period) |>
  kable()
```
@tbl-movie-release highlights the average number of movies released per year across the three periods. The pre-COVID period recorded the highest average, with approximately 263 movies released annually. During COVID, the average dropped significantly to around 167 movies per year, reflecting production disruptions. Post-COVID releases partially recovered, increasing to an average of 188 movies annually.


# Conclusion and Recommendations

