---
title: ""
author:
  - "Aryaman Dhindsa"
  - "Joshua Bannon"
  - "Smit Tambat"
  - "Chunjian Zhang"
course: "ETC5513"
unit: "Collaborative and Reproducible Practices"
semester: "Semester 1, 2026"
date: "27 May 2026"
format: html
self-contained: TRUE
---

# Executive Summary

# Introduction

When COVID-19 forced much of the world indoors, streaming quickly became one of the few forms of entertainment people could still rely on. Netflix, already the largest subscription service, was an obvious beneficiary of all that extra time spent at home. What is less obvious is whether all those new viewers were actually any happier with what they were watching. This report looks at that question through audience vote ratings, asking whether the average rating of Netflix films shifted before, during, and after the pandemic. We split the data into three periods which are pre-COVID (2010–2019), the COVID years (2020–2021), and post-COVID (2022–2024) and compare the average ratings across them. We focus on ratings rather than the number of films released, since the dataset holds a fixed number of titles per year and so cannot tell us anything meaningful about output. Specifically, we ask: **Did the average audience vote rating of Netflix films change before, during, and after COVID-19?**

This report examines how the production budgets and box-office revenues of Netflix films changed before, during, and after COVID-19 period. We define three periods: pre-COVID (2016-2019), COVID (2020-2021), and post-COVID (2022-2023). Using a public dataset of Netflix films titles, we compare yearly means of budget and revenue across these periods to identify any pandemic related shifts. The analysis focuses on financial measures rather than release volume, as the dataset is balanced by release year and therefore unsuitable for measuring output. Understanding these patterns offers insight into how the pandemic affected the economics of Netflix film production. Specifically, we ask: **How did the Budgets and Revenues of Netflix films change before, during, and after COVID-19?** 

# Methodology
The analysis draws on a public dataset of 16,000 Netflix film titles, each recorded with its release year and an average audience vote score. We dropped the 2025 titles before doing anything else. Those films had only just been released, most had barely been rated, and leaving them in pulled the averages down to numbers that clearly weren't real. That left 15,000 films spanning 2010 to 2024 to work with.

To get at the pandemic question, we sorted the films into three periods:
pre-COVID (2010–2019), COVID (2020–2021), and post-COVID (2022–2024). For each period we worked out the mean vote average, which gives us a simple way to compare how well-received films were across the three windows. We also calculated the average rating year by year, so we could see the smaller movements that a three-way split tends to hide. Everything was done in R using the `tidyverse` packages, and the code sits in the same Quarto document as this report so the whole thing can be reproduced from the raw data.

@fig-yearly plots the yearly average rating from 2010 to 2024, with the two pandemic years marked so any shift around them is easy to spot. @tbl-periods then pulls this together, showing the mean rating and the number of films in each period side by side. Between them, the figure gives the year-to-year detail and the table gives the bigger comparison.

```{r}
#| label: fig-yearly
#| fig-cap: "Yearly mean budget and revenue for Netflix films, 2016–2023 (USD)."
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

netflix <- read_csv("data/netflix_data.csv") |>
  filter(release_year >= 2016, release_year <= 2023,
         budget > 0, revenue > 0)

yearly <- netflix |>
  group_by(release_year) |>
  summarise(
    `Mean budget` = mean(budget, na.rm = TRUE),
    `Mean revenue` = mean(revenue, na.rm = TRUE)
  ) |>
  pivot_longer(cols = c(`Mean budget`, `Mean revenue`),
               names_to = "Measure", values_to = "USD")

ggplot(yearly, aes(x = release_year, y = USD, colour = Measure)) +
  geom_line(linewidth = 1) +
  geom_point() +
  geom_vline(xintercept = c(2019, 2021), linetype = "dashed",
             colour = "grey50") +
  scale_y_continuous(labels = scales::label_dollar(scale = 1e-6,
                                                    suffix = "M")) +
  scale_colour_manual(values = c("Mean budget" = "#221F1F",
                                  "Mean revenue" = "#E50914")) +
  labs(x = "Release year", y = "Mean (USD)", colour = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| label: tbl-periods
#| tbl-cap: "Mean budget and revenue by COVID-19 period (USD millions)."
#| tbl-cap-location: bottom
#| echo: false
#| message: false
#| warning: false

library(knitr)

netflix |>
  mutate(period = case_when(
    release_year <= 2019 ~ "Pre-COVID (2016–2019)",
    release_year <= 2021 ~ "COVID (2020–2021)",
    TRUE                 ~ "Post-COVID (2022–2023)"
  )) |>
  group_by(period) |>
  summarise(
    `Mean budget (USD M)` = round(mean(budget, na.rm = TRUE) / 1e6, 1),
    `Mean revenue (USD M)` = round(mean(revenue, na.rm = TRUE) / 1e6, 1),
    `Number of films` = n()
  ) |>
  arrange(factor(period, levels = c(
    "Pre-COVID (2016–2019)", "COVID (2020–2021)", "Post-COVID (2022–2023)"
  ))) |>
  rename(Period = period) |>
  kable()
```



# Result


# Conclusion and Recommendations
